---
title: "<center>Extracting raster values with 'unique' polygons</center>"
author: "<center>Wyclife Agumba Oluoch (wyclifeoluoch@gmail.com)</center>"
date: "<center>`r Sys.time()`</center>"
bibliography: bib/packages.bib
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
nocite: '@*'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::write_bib(c(.packages(),
                   'base',
                   'raster',
                   'stars',
                   'rgdal'), 'bib/packages.bib')
```
# Introduction

In this short article, I demonstrate results of extracting raster values with polygons which:

1. Touches several pixels but not any one completely covered.
2. Completely inside a single pixel but not covering centroid of the pixel.
3. Completely outside the raster pixels bounds.
4. Partially touches a pixel of the raster
5. Covers most part of a pixel but not centroid.
6. Covers least part of a pixel and centroid
7.

# Loading the libraries

To do this, I will need several libraries including:

a. base
b. raster
c. stars
d. rgdal

```{r load_libraries, include=FALSE}
library(rgdal)
library(raster)
library(stars)

```
# Loading the datasets

I use one raster layer and seven different polygon layers as stated in the introduction section..

```{r load_data, include=FALSE}
raster_data <- raster('data/raster.tif')
polygon <- readOGR('data/polygon.shp')
polygon2 <- readOGR('data/polygon2.shp')
polygon3 <- readOGR('data/polygon3.shp')
polygon4 <- readOGR('data/polygon4.shp')
polygon5 <- readOGR('data/polygon5.shp')
polygon6 <- readOGR('data/polygon6.shp')
polygon7 <- readOGR('data/polygon7.shp')
```

# Visualize pixel values

```{r pixel_values}
stars_data <- read_stars('data/raster.tif')
plot(stars_data, text_values = TRUE)
```

# Visualizing the 'unique' polygons on the raster

The seven unique polygons are overlayed on the raster image below to show their uniqueness in spatial analyses environment.

```{r warning = FALSE, echo=FALSE}
plot(raster_data)
plot(polygon, add = TRUE)
plot(polygon2, add = TRUE)
plot(polygon3, add = TRUE)
plot(polygon4, add = TRUE)
plot(polygon5, add = TRUE)
plot(polygon6, add = TRUE)
plot(polygon7, add = TRUE)
```


# References

